\HeaderA{Rserve}{Server providing R functionality to applications via TCP/IP or local unix sockets}{Rserve}
\keyword{interface}{Rserve}
\begin{Description}\relax
Starts Rserve in daemon mode (unix only). Any additional parameters
not related to Rserve will be passed straight to the underlying R. For
configuration, usage and command line parameters please consult the
online documentation at http://www.rosuda.org/Rserve. Use \code{R CMD Rserve
  --help} for a brief help.

The \code{Rserve} function is provided for convenience only, it is
recommended to start Rserve directly from the command line, not from R
itself. Also note that the debug version of Rserve doesn't fork and thus
will block until closed.
\end{Description}
\begin{Usage}
\begin{verbatim}
# R CMD Rserve [<parameters>]

Rserve(debug = FALSE, port = 6311, args = NULL)
\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{debug}] determines whether regular Rserve or debug version of
Rserve (\code{Rserve.dbg}) should be started.
\item[\code{port}] port used by Rserve to listen for connections
\item[\code{args}] further arguments passed to Rserve (as a string that will be
passed to the \code{system} command thus use quotes where necessary).
\end{ldescription}
\end{Arguments}
\begin{Details}\relax
Rserve is not just a package, but an application. It is provided as a
R package for convenience only. For details see
http://www.rosuda.org/Rserve
\end{Details}
\begin{Author}\relax
Simon Urbanek
\end{Author}

